comgen-outline

### 2017-07-22 paper/appendix-cueValidity.Rmd
clean up most of cue validity appendix

In considering generalizations about categories, or *generic language*, theories try to formalize the truth conditions by analogy to quantified statements (e.g., *some*, *most*, ...).
Here, the prevalence of the feature, or $P(f \mid k)$, is the critical quantity.
Quantifiers can be described as conditions on prevalence: 
$$
\denote{some} = \{P(f \mid k) > 0\}, \denote{most} = \{P(f \mid k) > 0.5\}, \denote{all} = \{P(f \mid k) = 1\}
$$
The question for formal semantic theories has been: is there some threshold $\theta$ beyond which generic statements become true?
Hard-constraints on $\theta$ are difficult to defend because of cases like *Mosquitos carry malaria*, where only the weakest threshold would suffice (i.e., *Some mosquitos carry malaria*).

@Leslie2007 notes that mosquitos are the only species to carry malaria, and perhaps this explains why the generic is true.
More formally, the inverse probability --- the probability of the kind $k$ given the feature $f$: $P(k \mid f)$ --- might be able to save statistical-based approaches [@Leslie2007; @Khemlani2012].
This inverse probability, referred to as *cue validity*, encodes the diagnosticity of the feature for the kind.
For example, after learning that an entity *carries malaria*, the probability that the entity is a *mosquito* is very high.

@@

\begin{enumerate}
\item{Completely diagnostic features: We contend there are a number of features that only one (or very small) of categories contain. Examples include: \emph{carrying malaria} (mosquitos), \emph{carrying Lyme disease} (ticks), \emph{having manes} (lions), \emph{having pouches} (kangaroos). The cue validity of these features for the corresponding categories should be very high (close to 1).}
\item{Completely absent features: For features that are completely absent in a kind, the cue validity should be extremely low or 0. There are many examples of these kinds of pairings; for example, \emph{has wings} (leopard), \emph{has a mane} (shark), "has spots" (kangaroo), "has a pouch" (tiger). }
\item{Completely undiagnostic features: We contend there are a number of features that almost every category contains. Examples include: \emph{is female} (robin), \emph{is male} (lion), \emph{is juvenile} (kangaroo), \emph{is full-grown} (leopard). The cue validity of these features for particular categories should be extremely low or 0. Learning that an entity is female tells you almost nothing about what kind of animal it is.}
\end{enumerate}

We collected cue validity ratings by running both a direct question and a free production experiment.
For the free production experiment, the cue validity is the proportion of responses of the target category (e.g., "mosquitos") for the property (e.g., "carries malaria").
Of primary interest is the measurement for the boundary conditions described above.
The experiments can be seen on \url{https://mhtess.github.io}.

@@

Participants completed the thirty target trials (corresponding to the thirty generic statements used in Expt. X) in addition to ten filler trials (total number of trials = 40).
The filler trials were made up of random category -- property pairings.
All trials were presented in a randomized order.

@@

@@

To calculate confidence intervals for the free production data, we resampled participants (with replacement) and computed the proportion of responses that were of the target category (e.g., the proportion of "mosquito" responses for the cue "carries malaria"). We did this one-thousand times to took the empirical 95\% quantile of those proportions.

@@

The measures deviate the most in their treatment of the undiagnostic features. 
Learning that an entity is female does not strongly imply that it is a robin, which is accurately reflected in the free production measure but not in the direct question measure.
A similar phenomena can be observed in direct question measure for the other undiagnostic features.

@@

Here, we treat each entry (participant free production or experimentally supplied category like mosquitos) as contributing to the prevalence prior. This will have the prevalence prior favor kinds that are easy to produce (like dogs and cats) as well as the experimentally supplied kinds (like mosquitos and robins).

@@

  }
  print(p)
}
# save(wp.cueValidity, file =
#        paste(project.path, model.path,
#              "results/prevPriorDerivedCueValidity.Rdata", sep = ""))
# load(file = 
#        paste(project.path, model.path, 
#              "results/prevPriorDerivedCueValidity.Rdata", sep = ""))
```

@@

```{r cvModel-uniformPrior}
cueValidityModel2 <- '
var cueValidity = function(){
  var kind = uniformDraw(data.allKinds);
  var responses = _.filter(data.prior, {Category: kind});
  var response = _.isEmpty(responses) ? {prevalence: 0} : uniformDraw(responses);
  observe(Bernoulli({p: response.prevalence/100}), true)
  return kind == data.kind[0]
}

>>>>

In considering generalizations about categories, or *generic language*, theories try to formalize the truth conditions by analogy to quantified statements (e.g., *some*, *most*, ...).
Here, the prevalence of the feature, or $P(f \mid k)$, is the critical quantity.
Quantifiers can be described as conditions on prevalence: 
$$
\denote{some} = \{P(f \mid k) > 0\}, \denote{most} = \{P(f \mid k) > 0.5\}, \denote{all} = \{P(f \mid k) = 1\}
$$
The driving question for formal semantic theories has been: is there some threshold $\theta$ beyond which generic statements become true?
Fixed constraints on $\theta$ (e.g., $\theta = 0.5$) are difficult to defend because of cases like *Mosquitos carry malaria*, where only the weakest threshold would suffice (i.e., *Some mosquitos carry malaria*).

With respect to malaria-carrying mosquitos, @Leslie2007 notes that the mosquito is the only species to carry malaria. 
The distinctiveness of *carrying malaria* then perhaps why *Mosquitos carry malaria* is true.
More formally, the inverse probability --- the probability of the kind $k$ given the feature $f$: $P(k \mid f)$ --- might be able to save statistical-based approaches [@Leslie2007; @Khemlani2012].
This inverse probability, referred to as *cue validity*, encodes the diagnosticity of the feature for the kind.
For example, after learning that an entity *carries malaria*, the probability that the entity is a *mosquito* is very high.

@@

\begin{enumerate}
\item{Completely diagnostic features: We contend there are a number of features that only one (or very small) of categories contain. Examples include: \emph{carrying malaria} (mosquitos), \emph{carrying Lyme disease} (ticks), \emph{having manes} (lions), \emph{having pouches} (kangaroos). The cue validity of these features for the corresponding categories should be very high (close to 1).}
\item{Completely absent features: For features that are completely absent in a kind, the cue validity should be extremely low or 0. There are many examples of these kinds of pairings; for example, \emph{has wings} (leopard), \emph{has a mane} (shark), *has spots* (kangaroo), *has a pouch* (tiger). }
\item{Completely undiagnostic features: We contend there are a number of features that almost every category contains. Examples include: \emph{is female} (robin), \emph{is male} (lion), \emph{is juvenile} (kangaroo), \emph{is full-grown} (leopard). The cue validity of these features for particular categories should be extremely low or 0. Learning that an entity is female tells you almost nothing about what kind of animal it is.}
\end{enumerate}

We collected cue validity ratings by running both a direct question and a free production experiment.
For the free production experiment, the cue validity is the proportion of responses of the target category (e.g., "mosquitos") for the property (e.g., "carries malaria").
Of primary interest is the measurement for the boundary conditions described above.
Links to the experiments can be found on \url{https://mhtess.github.io}.

@@

Participants completed the thirty target trials (corresponding to the thirty generic statements used in Expt. 1b) in addition to ten filler trials (total number of trials = 40).
The filler trials were made up of random category -- property pairings.
All trials were presented in a randomized order.

@@

### Free production data processing

To calculate confidence intervals for the free production data, we resampled participants (with replacement) and computed the proportion of responses that were of the target category (e.g., the proportion of "mosquito" responses for the cue "carries malaria"). We did this 1000 times to generate an empirical distribution from which 95\% intervals could be calculated.

@@

@@

The measures deviate in their characterization of the undiagnostic features. 
Learning that an entity is female does not strongly imply that it is a robin, which is accurately reflected in the free production measure but not in the direct question measure.
A similar phenomenon can be observed in direct question measure for the other undiagnostic features.
Figure \ref{fig:cv-bothquestions-barplots} (bottom) shows the raw empirical distributions of responses for the direct question measure. 
We observe suggestive evidence that participants respond to this question for unidagnostic features in one of two ways: (i) reporting near-0 likelihood (accurate response) or (ii) reported near-0.5 likelihood. 
This latter response option may be a pragmatic effect of asking the direct question and/or participants "opting out" of a response. 
For example, in response to the question "There is a thing that is female. What are the odds that it is a robin?", a reasonable human response may be "I don't know". Participants may cope with the awkwardness of the question by placing the slider bar in the middle of the scale.

@@

We can further understand our measures of cue validity by comparing them to cue validity derived from our prevalence prior elicitation task (Expt. 1a).
Expt. 1a is not perfectly designed for this comparison, as we supplied participants with half of the animal categories that they rated. (Participants freely generated the other half). Thus, we do not have the best measurement of the prior probability of categories $P(k)$.

In this analysis, we treat each category entry from Expt. 1a (participant free production or experimentally supplied category like mosquitos) as contributing to the prevalence prior.
This will result in the prevalence prior favoring kinds that are easy to produce (like dogs and cats, which is plausibly a good approximation for $P(k)$) as well as favoring the experimentally supplied kinds (like mosquitos and robins).

We compute cue validity from the prevalence prior using Bayes' rule. 
Figure \ref{fig:cv-scatters} show the ...

@@

  }
  #print(p)
}
# save(wp.cueValidity, file =
#        paste(project.path, model.path,
#              "results/prevPriorDerivedCueValidity.Rdata", sep = ""))
# load(file = 
#        paste(project.path, model.path, 
#              "results/prevPriorDerivedCueValidity.Rdata", sep = ""))
```

@@

```{r cvModel-uniformPrior, eval = F}
cueValidityModel2 <- '
var cueValidity = function(){
  var kind = uniformDraw(data.allKinds);
  var responses = _.filter(data.prior, {Category: kind});
  var response = _.isEmpty(responses) ? {prevalence: 0} : uniformDraw(responses);
  observe(Bernoulli({p: response.prevalence/100}), true)
  return kind == data.kind[0]
}
